---
title: "HW6"
output: html_notebook
---

```{r}
getwd()
```

Connect the database
```{r}
library(RSQLite)
fpath=""
dbfile="homeworkDB.sqlite"
dbcon<-dbConnect(RSQLite::SQLite(), paste0(fpath, dbfile))
```

```{r} 
dbExecute(dbcon, "PRAGMA foreign_keys = ON")
```


```{sql connection=dbcon}
DROP TABLE IF EXISTS affiliations
```

Create a lookup table for affiliation
```{sql connection=dbcon}
CREATE TABLE IF NOT EXISTS affiliations (
  affid INTEGER NOT NULL PRIMARY KEY,
  affname TEXT NOT NULL
)
```

Insert values into lookup table
```{sql connection=dbcon}
INSERT INTO affiliations VALUES 
  (1, "PT"),
  (2, "FT"),
  (3, "Contract")
```

```{sql connection=dbcon}
DROP TABLE IF EXISTS authors
```

Create an Author table
```{sql connection=dbcon}
CREATE TABLE IF NOT EXISTS authors (
  aid INTEGER PRIMARY KEY AUTOINCREMENT,
  name TEXT NOT NULL,
  email TEXT NOT NULL,
  affiliation TEXT NOT NULL,
  FOREIGN KEY (affiliation) REFERENCES affiliations (affid)
)
```

Insert values into Author table
```{sql connection=dbcon}
INSERT INTO authors(name, email, affiliation) VALUES
  ("Jeff Lohan", "JeffLohan@gmail.com", 3),
  ("Alison Chu", "AlisonChu@gmail.com", 1),
  ("Lily Chen", "LilyChen@gmail.com", 2)
```

Check the Author table
```{sql connection=dbcon}
SELECT m.aid AS 'ID', m.name AS 'Name', m.email AS 'Email', r.affname AS 'AFF'
  FROM authors m INNER JOIN affiliations r ON (m.affiliation=r.affid);
```

Create log table
```{sql connection=dbcon}
CREATE TABLE authorsAuditLog(
  id INTEGER PRIMARY KEY AUTOINCREMENT,
  DmlTimestamp datetime NOT NULL,
  DmlType varchar(10) NOT NULL CHECK (DmlType IN ('INSERT', 'UPDATE', 'DELETE')),
  OldRowData nvarchar(1000),
  NewRowData nvarchar(1000)
)
```

Create delete trigger
```{sql connection=dbcon}
CREATE TRIGGER on_authors_delete_trigger 
AFTER DELETE
ON authors
BEGIN
  INSERT INTO authorsAuditLog(
    DmlType,
    OldRowData,
    NewRowData,
    DmlTimestamp
  ) 
  values(
    "DELETE",
    ("{Name:" || old.name || " Email:" || old.email || " Affiliation:" ||
      (SELECT r.affname FROM affiliations r WHERE (old.affiliation =r.affid)) || "}"
    ),
    NULL,
    DATETIME('NOW')
  );
END;
```

Original Author tabel
```{sql connection=dbcon}
SELECT m.aid AS 'ID', m.name AS 'Name', m.email AS 'Email', r.affname AS 'AFF'
  FROM authors m INNER JOIN affiliations r ON (m.affiliation=r.affid);
```

Delete one authors
```{sql connection=dbcon}
DELETE FROM authors WHERE aid=1
```

New Authors table
```{sql connection=dbcon}
SELECT m.aid AS 'ID', m.name AS 'Name', m.email AS 'Email', r.affname AS 'AFF'
  FROM authors m INNER JOIN affiliations r ON (m.affiliation=r.affid);
```
A 'DELETE' record is added into authorsAuditLog table
```{sql connection=dbcon}
SELECT * FROM authorsAuditLog;
```
Create insert trigger
```{sql connection=dbcon}
CREATE TRIGGER on_authors_insert_trigger 
BEFORE INSERT
ON authors
BEGIN
  INSERT INTO authorsAuditLog(
    DmlType,
    OldRowData,
    NewRowData,
    DmlTimestamp
  ) 
  values(
    "INSERT",
    NULL,
    ("{Name:" || new.name || " Email:" || new.email || " Affiliation:" ||
      (SELECT r.affname FROM affiliations r WHERE (new.affiliation =r.affid)) || "}"
    ),
    DATETIME('NOW')
  );
END;
```

Insert an author record
```{sql connection=dbcon}
INSERT INTO authors(name, email, affiliation) VALUES
  ("Lucy Zhou", "LucyZhou@gmail.com", 1)
```

New Authors table
```{sql connection=dbcon}
SELECT m.aid AS 'ID', m.name AS 'Name', m.email AS 'Email', r.affname AS 'AFF'
  FROM authors m INNER JOIN affiliations r ON (m.affiliation=r.affid);
```
A 'INSERT' record is added into authorsAuditLog table
```{sql connection=dbcon}
SELECT * FROM authorsAuditLog;
```
Create update trigger
```{sql connection=dbcon}
CREATE TRIGGER on_authors_update_trigger 
AFTER UPDATE
ON authors
BEGIN
  INSERT INTO authorsAuditLog(
    DmlType,
    OldRowData,
    NewRowData,
    DmlTimestamp
  ) 
  values(
    "UPDATE",
    ("{Name:" || old.name || " Email:" || old.email || " Affiliation:" ||
      (SELECT r.affname FROM affiliations r WHERE (old.affiliation =r.affid)) || "}"
    ),
    ("{Name:" || new.name || " Email:" || new.email || " Affiliation:" ||
      (SELECT r.affname FROM affiliations r WHERE (new.affiliation =r.affid)) || "}"
    ),
    DATETIME('NOW')
  );
END;
```

Update one author
```{sql connection=dbcon}
UPDATE authors SET name="NewName" WHERE name="Alison Chu"
```

New authors table
```{sql connection=dbcon}
SELECT m.aid AS 'ID', m.name AS 'Name', m.email AS 'Email', r.affname AS 'AFF'
  FROM authors m INNER JOIN affiliations r ON (m.affiliation=r.affid);
```

A 'UPDATE' record is added into authorsAuditLog table
```{sql connection=dbcon}
SELECT * FROM authorsAuditLog;
```

Create emailCheck trigger
```{sql connection=dbcon}
CREATE TRIGGER on_authors_emailCheck_trigger 
BEFORE 
INSERT 
ON authors
FOR EACH ROW
BEGIN
  SELECT
      CASE
  -- Check if the new email address is in the correct format
	      WHEN NEW.email NOT LIKE '%_@_%._%' THEN
	 -- If the email address is not in the correct format, raise an error
   	      RAISE (ABORT,'Invalid email address')
       END;
END;
)
```


Insert a row with an invalid email into the authors table.
SQLite issued an error: “Invalid email address” and aborted the execution of the insert.
```{sql connection=dbcon}
INSERT INTO authors(name, email, affiliation) VALUES
  ("Jeff Lohan", "JeffLohangmail.com", 3)
```

Insert a row with a valid email.
```{sql connection=dbcon}
INSERT INTO authors(name, email, affiliation) VALUES
  ("Jack Wang", "JackWang@gmail.com", 1)
```

Because the email is valid, the insert statement executed successfully.
```{sql connection=dbcon}
SELECT m.aid AS 'ID', m.name AS 'Name', m.email AS 'Email', r.affname AS 'AFF'
  FROM authors m INNER JOIN affiliations r ON (m.affiliation=r.affid);
```


```{sql connection=dbcon}
DROP TABLE IF EXISTS modules
```

Create a Module table
```{sql connection=dbcon}
CREATE TABLE IF NOT EXISTS modules (
  number TEXT NOT NULL PRIMARY KEY,
  title TEXT NOT NULL,
  lengthInMin INTEGER NOT NULL
)
```

Insert values into Module table
```{sql connection=dbcon}
INSERT INTO modules(number, title, lengthInMin) VALUES
  ("A","math", "50"),
  ("B", "database", "100")
```

Test module table
```{sql connection=dbcon}
SELECT * FROM modules;
```

```{sql connection=dbcon}
DROP TABLE IF EXISTS AuthorModule
```

Create a AuthorModule/junction table
```{sql connection=dbcon}
CREATE TABLE IF NOT EXISTS AuthorModule
(
    aid INTEGER NOT NULL,
    number TEXT,
    CONSTRAINT PK_AuthorModule PRIMARY KEY
    (
        aid,
        number
    ),
    FOREIGN KEY (number) REFERENCES modules (number),
    FOREIGN KEY (aid) REFERENCES authors (aid)
)
```

Disconnect the database
```{r}
dbDisconnect(dbcon)
```

