---
title: "R Notebook"
output: html_notebook
---

connect to local MySQL database
```{r}
# Load the packages
library(RMySQL)

db_user <- 'root'
db_password <- 'Qazwsx123.'
db_name <- 'practicum1'
db_host <- '127.0.0.1' 
db_port <- 3306
mydb <-  dbConnect(MySQL(), user = db_user, password = db_password,
                 dbname = db_name, host = db_host, port = db_port)
```

```{r} 
dbExecute(mydb, "SET FOREIGN_KEY_CHECKS=1;")
```

```{sql connection=mydb}
DROP TABLE IF EXISTS conditions
```

Create a lookup table conditions
```{sql connection=mydb}
CREATE TABLE IF NOT EXISTS conditions (
  cid INTEGER NOT NULL PRIMARY KEY,
  condition_name VARCHAR(50) NOT NULL,
  explanation VARCHAR(255)
)
```

Insert values into lookup table conditions
```{sql connection=mydb}
INSERT INTO conditions VALUES 
  (1, "No Cloud", ""),
  (2, "Some Cloud", ""),
  (3, "Overcast", "")
```


```{sql connection=mydb}
DROP TABLE IF EXISTS airports
```

Create a table that stores airports and states called airports
The value of airportName is the data of airport column in csv table.
The value of state is the data of origin column in csv table.
```{sql connection=mydb}
CREATE TABLE IF NOT EXISTS airports (
  aid INTEGER NOT NULL AUTO_INCREMENT PRIMARY KEY ,
  airportName TEXT NOT NULL,
  state TEXT NOT NULL,
  airportCode TEXT
)
```


```{sql connection=mydb}
DROP TABLE IF EXISTS incidents
```

Create a table that stores wildlife strike incidents called incidents
Assuming that the incidents table has a column named "origin" 
that references the "aid" column in the airports table
```{sql connection=mydb}
CREATE TABLE IF NOT EXISTS incidents (
  rid INTEGER NOT NULL PRIMARY KEY ,
  date DATE,
  origin INTEGER NOT NULL,
  airline VARCHAR(255) NOT NULL,
  aircraft VARCHAR(255) NOT NULL,
  flightPhase VARCHAR(255) 
      CHECK (flightPhase= 'takeoff' OR flightPhase= 'landing' OR 
             flightPhase= 'inflight' OR flightPhase= 'unknown'),
  altitude VARCHAR(255) NOT NULL,
  conditions INTEGER NOT NULL,
  warning TINYINT NOT NULL,
  FOREIGN KEY (origin) REFERENCES airports (aid),
  FOREIGN KEY (conditions) REFERENCES conditions (cid)
)
```

Create  stored procedure to remove all military flights from the database.
```{sql connection=mydb}
CREATE PROCEDURE removeMilitaryFlights()
BEGIN
  DELETE FROM incidents WHERE airline = 'military';
END;

```


Firstly, Select only the columns("rid", "aircraft", "airport", "flight_date", 
"airline", "origin", "flight_phase", "sky_conditions", "pilot_warned_flag", 
"altitude_ft") that are needed from the csv table and stored in bird_strikes data frame.
Sencondly, remove all rows whose "rid" value in not an integer, "pilot_warned_flag" value
is not "Y" or "N", "sky_conditions" value is not "Overcast", "Some Cloud" or "No Cloud".
Then remove all rows with missing values in any column of data frame.
Finally, manipulate the remaining rows in the data frame.

```{r}
library(dplyr)
# Read in the CSV file
bird_strikes <- read.csv("BirdStrikesData-V2.csv", header = TRUE, stringsAsFactors = FALSE)

# Select only the columns that are needed
bird_strikes <- bird_strikes[c("rid", "aircraft", "airport", "flight_date", 
                               "airline", "origin", "flight_phase", "sky_conditions"
                               , "pilot_warned_flag", "altitude_ft")]


# Filter row whose values meet the criteria
bird_strikes <- bird_strikes %>% 
                filter(is.integer(rid),
                       pilot_warned_flag %in% c("Y", "N"),
                       sky_conditions %in% c("Overcast", "Some Cloud", "No Cloud"))

#Remove rows with missing values in any column of data frame
bird_strikes <- bird_strikes[complete.cases(bird_strikes), ]

# Loop through each row of the bird_strikes data frame
for (i in 1:nrow(bird_strikes)) {
  # Get the values fo "rid", "aircraft", "airport", "flight_date", 
  #"airline", "origin", "flight_phase", "sky_conditions", 
  # "pilot_warned_flag", "altitude_ft" for the current row
  rid_value <- bird_strikes$rid[i]
  airline_value <- bird_strikes$airline[i]
  aircraft_value <- bird_strikes$aircraft[i]
  altitude_ft_value <- bird_strikes$altitude_ft[i]
  
  
  #Code "Landing Roll" as "landing";
  #Code 'Approach' as 'landing'
  #code 'Climb' as 'takeoff'
  #code "Take-off run" as "takeoff"
  #code "Descent" as "inflight"
  #code Other flight phases as "unknown"
  #Harmonize the flight phases to be one of: takeoff, landing, inflight, unknown.
  flight_phase_value <- bird_strikes$flight_phase[i]
  if((flight_phase_value=="Landing Roll") | (flight_phase_value=="Approach")){
    flight_phase_obj<-"landing"
  }else if((flight_phase_value=="Climb") | (flight_phase_value=="Take-off run")){
    flight_phase_obj<-"takeoff"
  }else if(flight_phase_value=="Descent"){
    flight_phase_obj<-"inflight"
  }else{
    flight_phase_obj<-"unknown"
  }
  
  #Make 'warning' a Boolean flag and use TRUE if the pilot was warned, FALSE otherwise.
  pilot_warned_flag_value <- bird_strikes$pilot_warned_flag[i]
  if(pilot_warned_flag_value=="Y"){
    warning_obj <- TRUE;
  }else if(pilot_warned_flag_value=="N"){
    warning_obj <- FALSE;
  }
  
  #Only store the date of the incident and store the date as a date type.
  flight_date_value <- strsplit(bird_strikes$flight_date[i], " ")[[1]][1]
  date_obj <- strptime(flight_date_value, '%m/%d/%Y')
  
  #Escape the single quote character by adding another single quote before it
  # to prevent errors when running SQL queries
  airport_value <- bird_strikes$airport[i]
  airport_value <- gsub("'", "''", airport_value)
  
  # Check if the origin is already in the airports table
  origin_value <- bird_strikes$origin[i]
  origin_query <- sprintf("SELECT aid FROM airports 
                          WHERE airportName = '%s' AND state = '%s'", 
                          airport_value, origin_value)
  origin_result <- dbGetQuery(mydb, origin_query)
  
  
  # If the airport is not in the airports table, insert it and get the aid
  if (nrow(origin_result) == 0 ) {
    airport_query <- sprintf("INSERT INTO airports (airportName, state) 
                             VALUES ('%s', '%s')", airport_value, origin_value)
    dbSendQuery(mydb, airport_query)
    origin_aid <- dbGetQuery(mydb, "SELECT LAST_INSERT_ID()")$`LAST_INSERT_ID()`
  } else {
    origin_aid <- origin_result$aid[1]
  }
  
  # Prepare the SQL query to look up the cid value in the conditions table
  sky_conditions_value <- bird_strikes$sky_conditions[i]
  conditions_query <- sprintf("SELECT cid FROM conditions WHERE condition_name = '%s'", 
                              sky_conditions_value)
  
  # Execute the query and get the cid value according to lookup table conditions
  condition_cid <- dbGetQuery(mydb, conditions_query)$cid
  
  # Prepare the SQL query to insert data into the incidents table
  incidents_query <- sprintf("INSERT INTO incidents 
                              (rid, date, origin, airline, aircraft, flightPhase, altitude, conditions, warning)
                             VALUES (%d, '%s', %d, '%s', '%s', '%s', '%s', %d, %d)", 
                             rid_value, 
                             date_obj, 
                             origin_aid, 
                             airline_value,  
                             aircraft_value,
                             flight_phase_obj,
                             altitude_ft_value, 
                             condition_cid,
                             as.integer(warning_obj))
  
  # Insert data into the incidents table
  dbSendQuery(mydb, incidents_query)
  }

# Call the stored procedure to delete military airlines from the incidents table
delete_military_airlines <- dbSendQuery(mydb, "CALL removeMilitaryFlights()")
```

Q3:Show that the loading of the data worked by displaying parts of each table 
```{r}
# Show the first 10 rows of the airports table
airports_query <- "SELECT * FROM airports LIMIT 10"
airports_df <- dbGetQuery(mydb, airports_query)
print(airports_df)

# Show the first 10 rows of the incidents table
incidents_query <- "SELECT * FROM incidents LIMIT 10"
incidents_df <- dbGetQuery(mydb, incidents_query)
print(incidents_df)

# Show the conditions table
conditions_query <- "SELECT * FROM conditions"
conditions_df <- dbGetQuery(mydb, conditions_query)
print(conditions_df)
```

Q4: Find the top-10 airlines with the most number of incidents
```{sql connection=mydb}
SELECT airline, COUNT(*) AS num_incidents
FROM incidents
GROUP BY airline
ORDER BY num_incidents DESC
LIMIT 10;
```

Q5:find the flight phase that had an above average number bird strike incidents
```{sql connection=mydb}
SELECT flightPhase, COUNT(*) AS num_incidents
FROM incidents
GROUP BY flightPhase
HAVING COUNT(*) > (
  SELECT AVG(num_incidents) AS avg_num_incidents 
  FROM
  (
    SELECT flightPhase, COUNT(*) AS num_incidents
    FROM incidents
    GROUP BY flightPhase
  ) a
)

```

Q6:find the maximum number of bird strike incidents by month (across all years)
Find the sum of all incidents across all years for each month. 
Then sort to find the maximum incidents.
```{sql connection=mydb}
SELECT MONTH(date) AS month, COUNT(*) AS num_incidents
FROM incidents
GROUP BY month
HAVING num_incidents =(
  SELECT MAX(num_incidents)
  FROM 
  (
    SELECT MONTH(date) AS month, COUNT(*) AS num_incidents
    FROM incidents
    GROUP BY month
  ) a
)
```

Q7:Build a column chart that visualizes the number of 
bird strikes incidents per year from 2005 to 2011
```{r}
library(ggplot2)
# Create a SQL query that calculates the number of bird strike incidents per year from 2005 to 2011
sql_query <- "SELECT YEAR(date) AS year, COUNT(*) AS num_incidents
              FROM incidents
              WHERE YEAR(date) BETWEEN 2005 AND 2011
              GROUP BY year
              ORDER BY year ASC"

# Use dbGetQuery() to execute the SQL query and retrieve the results as a data frame
bird_strikes_by_year <- dbGetQuery(mydb, sql_query)

# Use ggplot2 to create a column chart
ggplot(bird_strikes_by_year, aes(x = year, y = num_incidents, fill = factor(year))) +
  geom_col() +
  scale_fill_manual(values = c("#1f77b4", "#ff7f0e", "#2ca02c", "#d62728", "#9467bd", "#8c564b", "#e377c2")) +
  labs(x = "Year", y = "Number of Incidents", title = "Bird Strikes Incidents per Year (2005-2011)") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14, face = "bold", hjust = 0.5),
        axis.title = element_text(size = 12, face = "bold"),
        axis.text = element_text(size = 10),
        legend.title = element_blank(),
        legend.text = element_text(size = 10))
```

Q8:
The parameters needed to be passed to the stored procedure:aircraft_value, airport_value,flight_year_value, flight_month_value, flight_day_value, airline_value, origin_value, flight_phase_value, sky_conditions_value, pilot_warned_flag_value,
altitude_ft_value. After call this stored procedure, if the given parameters meet 
requirements, new incident will be added into database.
```{sql connection=mydb}
CREATE PROCEDURE addNewIncident (
  IN aircraft_value VARCHAR(255),
  IN airport_value VARCHAR(255),
  IN flight_year_value INT,
  IN flight_month_value INT,
  IN flight_day_value INT,
  IN airline_value VARCHAR(255),
  IN origin_value VARCHAR(255),
  IN flight_phase_value VARCHAR(255),
  IN sky_conditions_value VARCHAR(255),
  IN pilot_warned_flag_value VARCHAR(255),
  IN altitude_ft_value VARCHAR(255)
)
BEGIN

  DECLARE new_cid INT;
  DECLARE airport_exists INT;
  DECLARE new_aid INT;
  DECLARE warning TINYINT;
  DECLARE max_rid INT DEFAULT 0;
  DECLARE flight_date DATE;
  DECLARE flight_phase VARCHAR(50);
  
 -- Check if flight_year_value\flight_month_value\flight_day_value are all reasonable values
  IF (flight_year_value < 1900 OR flight_year_value > YEAR(CURDATE()))  THEN
    SIGNAL SQLSTATE '45000' SET MESSAGE_TEXT = 'Invalid flight_year_value value';
  END IF;
  IF (flight_month_value < 1 OR flight_month_value > 12)  THEN
    SIGNAL SQLSTATE '45000' SET MESSAGE_TEXT = 'Invalid flight_month_value value';
  END IF;
  IF (flight_day_value < 1 OR flight_day_value > 31)  THEN
    SIGNAL SQLSTATE '45000' SET MESSAGE_TEXT = 'Invalid flight_day_value value';
  END IF;

  -- Check if "pilot_warned_flag" value if  "Y" or "N"
  IF pilot_warned_flag_value != "Y" AND pilot_warned_flag_value != "N" THEN
    SIGNAL SQLSTATE '45000' SET MESSAGE_TEXT = 'Invalid pilot_warned_flag value';
  END IF;
  
  -- Check if "sky_conditions" value is "Overcast", "Some Cloud" or "No Cloud"
  IF sky_conditions_value != "Overcast" 
      AND sky_conditions_value != "Some Cloud" 
      AND sky_conditions_value != "No Cloud" THEN
    SIGNAL SQLSTATE '45000' SET MESSAGE_TEXT = 'Invalid sky_conditions value';
  END IF;
 
 -- Get the cid number from conditions table according to sky_conditions_val
  SELECT cid INTO new_cid FROM conditions WHERE condition_name = sky_conditions_value;
  
  -- Check if airport is already in the airports table
  SET airport_exists =(SELECT COUNT(*) FROM airports 
                        WHERE airportName = airport_value AND state = origin_value);
  -- If airport is not in the airports table, insert it and get the new aid
  IF airport_exists = 0 THEN
    INSERT INTO airports (airportName, state) VALUES (airport_value, origin_value);
    SELECT LAST_INSERT_ID() INTO new_aid;
  ELSE
    SELECT aid INTO new_aid FROM airports 
               WHERE airportName = airport_value AND state = origin_value;
  END IF;
  
  -- Change pilot_warned_flag_value to the TRUE OR FALSE
  IF pilot_warned_flag_value = "Y" THEN
    SET warning = 1;
  ELSE
    SET warning = 0;
  END IF;
  
  -- Get the maximum rid from the incidents table and add 1 to produce the new rid
  SELECT MAX(rid) INTO max_rid FROM incidents;
  SET max_rid = max_rid + 1;
  
  -- Get the date object according to the given flight_year_value, flight_month_value, flight_day_value
  SET flight_date=STR_TO_DATE(CONCAT(flight_year_value, '-', flight_month_value, '-', flight_day_value),'%Y-%m-%d');
  
  -- Change the given flight_phase_value to "takeoff, landing, inflight, unknown".
  SET flight_phase = 
    CASE
      WHEN flight_phase_value = 'Landing Roll' OR flight_phase_value = 'Approach' THEN 'landing'
      WHEN flight_phase_value = 'Climb' OR flight_phase_value = 'Take-off run' THEN 'takeoff'
      WHEN flight_phase_value = 'Descent' THEN 'inflight'
      ELSE 'unknown'
    END;
  
  -- Insert the new incident into the incidents table
  INSERT INTO incidents (rid, date, origin, airline, aircraft, flightPhase, altitude, conditions, warning)
  VALUES (max_rid, flight_date, new_aid, airline_value, aircraft_value, flight_phase, 
  altitude_ft_value, new_cid, warning);
END
```

Call this stored procedure in R using the RMySQL package.
Test case 1: No new airport is added
```{r}
# Execute the stored procedure--airport already exits
originAirportsNumber <- dbGetQuery(mydb, "SELECT COUNT(*) FROM airports")

query <- paste("CALL addNewIncident('TEST1', 'JOHN F KENNEDY INTL'
,2022,3,5,'Delta Air Lines','New York','Landing Roll','Some Cloud', 'Y', '3000')")
result <- dbSendQuery(mydb, query)

# check if the new record was successfully added
if (dbGetRowsAffected(result) == 1) {
  message('New incident added successfully')
} else {
  warning('Failed to add new incident')
}

# Execute a SELECT statement to retrieve the new record
result <- dbGetQuery(mydb, "SELECT * FROM incidents where aircraft='TEST1'")

#Pring new incident record.
print(result)

# Execute a SELECT statement to retrieve the new record
newAirportsNumber <- dbGetQuery(mydb, "SELECT COUNT(*) FROM airports")

# Check if the result is not empty, indicating that the new record was successfully added
if(newAirportsNumber - originAirportsNumber == 1) {
  print("New airport was successfully added to the airports table")
} else if(newAirportsNumber == originAirportsNumber){
  print("No new airport was added to the airports table")
}
```

Test case 2: New airport is added
```{r}
# Execute the stored procedure--new airport is added.
originAirportsNumber <- dbGetQuery(mydb, "SELECT COUNT(*) FROM airports")

query <- paste("CALL addNewIncident('TEST2', 'TEST AIRPORT'
,2022,3,6,'Delta Air Lines','New York','Landing Roll','No Cloud', 'Y', '5000')")
result <- dbSendQuery(mydb, query)

# check if the new record was successfully added
if (dbGetRowsAffected(result) == 1) {
  message('New incident added successfully')
} else {
  warning('Failed to add new incident')
}

# Execute a SELECT statement to retrieve the new record
newIncident <- dbGetQuery(mydb, "SELECT * FROM incidents where aircraft='TEST2'")

#Pring new incident record.
print(newIncident)

# Execute a SELECT statement to retrieve the new airport record
newAirport <- dbGetQuery(mydb, 
            "SELECT * FROM airports where airportName='TEST AIRPORT' AND state='New York'")

#Pring new airport record.
print(newAirport)

# Execute a SELECT statement to retrieve the new record
newAirportsNumber <- dbGetQuery(mydb, "SELECT COUNT(*) FROM airports")

# Check if the result is not empty, indicating that the new record was successfully added
if(newAirportsNumber - originAirportsNumber == 1) {
  print("New airport was successfully added to the airports table")
} else if(newAirportsNumber == originAirportsNumber){
  print("No new airport was added to the airports table")
}
```


```{r}
dbDisconnect(mydb)
```
