---
title: "R Notebook"
output: html_notebook
---

```{r}
getwd()
```

Connect the database
```{r}
library(RSQLite)
fpath=""
dbfile="MediaDB.db"
dbcon<-dbConnect(RSQLite::SQLite(), paste0(fpath, dbfile))
```


Q1:The last names, city, and state of all customer who made at least one purchase
```{sql connection=dbcon}
select c.LastName, c.City, c.State 
from customers c inner join invoices i 
on c.CustomerId=i.CustomerId
group by c.CustomerId
```

Q2:The titles of each albums and the number of tracks on each album
```{sql connection=dbcon}
select a.Title, count(t.TrackId) as No_Of_Tracks
from tracks t inner join albums a on t.AlbumId=a.AlbumId
group by a.Title
```

Q3: List the number of tracks by media type
```{sql connection=dbcon}
select m.Name as 'Media type', count(t.TrackId) as No_Of_Tracks
from media_types m left join tracks t on m.MediaTypeId=t.MediaTypeId
group by m.Name
```


Q4:The average age of the employees
```{sql connection=dbcon}
select avg((strftime('%s','now') - strftime('%s', BirthDate))/31536000) as 'Average age'
from employees
```

Q5:The total number of unique customers for each US state
ordered alphabetically by state, who made at least one purchase
```{sql connection=dbcon}
select c.State as 'State', count(distinct c.CustomerId) as No_of_unique_customers
from customers c inner join invoices i on c.CustomerId=i.CustomerId
where 	c.Country='USA'
group by c.State
order by c.State
```

Q6:Customers never made a purchase
```{sql connection=dbcon}
select count(*) as No_Of_noPurchase_customer
from customers c left join invoices i on c.CustomerId=i.CustomerId
where i.InvoiceId is null
```

Q7:The number of albums contain the substring "symphony" in the album title
```{sql connection=dbcon}
select count(*) as No_Of_Albums from albums
where Title like '%symphony%';
```

Q8:The total "time" of all tracks per artist
```{sql connection=dbcon}
select art.Name as 'Artist', round(sum(t.Milliseconds)/1000.0/3600, 2) as Time_in_hour
from albums a
inner join tracks t on a.AlbumId=t.AlbumId
inner join artists art on art.ArtistId=a.ArtistId
group by art.Name
```

Q9:Which artist(s) published the most tracks, regardless of type and format? List the artist(s) name(s).
```{sql connection=dbcon}
select art.Name as 'Artist', count(t.TrackId) as No_Of_Track
from albums a
inner join tracks t on a.AlbumId=t.AlbumId
inner join artists art on art.ArtistId=a.ArtistId
group by art.Name
order by No_Of_Track DESC
limit 1
```

Q10:Which artists have composed tracks in more than three different genres? List the artists name and the number of distinct genres they have compositions in.
```{sql connection=dbcon}
select art.Name as 'Artist', count(distinct g.GenreId) as No_Of_distinct_genres
from albums a
inner join tracks t on a.AlbumId=t.AlbumId
inner join artists art on art.ArtistId=a.ArtistId
inner join genres g on t.GenreId=g.GenreId
group by art.Name
having count(distinct g.GenreId)>3
```

Disconnect the database
```{r}
dbDisconnect(dbcon)
```